##Coursera Project 1 - Evaluating Movement Data


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/danielmonaco1/Desktop")
```

First thing we are going to do is download the data.

```{r cache = TRUE}
data = read.csv("activity.csv")
```

# Question 1: What is mean total number of steps taken per day?

```{r Mean Data}
hist(data$steps)
mean = mean(data$steps,na.rm = TRUE)
median = median(data$steps,na.rm = TRUE)
```

So it looks like the mean steps taken per day was `r mean` and the median was `r median`.

# What is the average daily activity pattern?

```{r}
library(dplyr)
timeseries = group_by(data, interval) %>% summarize(m = mean(steps,na.rm = TRUE))
plot(timeseries$interval,timeseries$m,type = "l",xlab = "Interval",ylab = "Average Steps",main = "Analyzing Average Steps vs Interval Time")
max= timeseries$interval[which(timeseries$m == max(timeseries$m))]
```

From this we can tell the interval with the largest average steps is `r max`

# Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

```{r}
NAN= sum(is.na(data$steps))
NAN_data = data[is.na(data$steps),]
value = data %>% group_by(interval) %>% summarize(m = mean(steps,na.rm = TRUE))
for(R1 in 1:NAN){
  NAN_data[R1,1] = value[which(NAN_data[R1,3] == value[,1]),2]
}
data2 = rbind(data[!is.na(data$steps),],NAN_data)

hist(data2$steps)
mean2 = mean(data2$steps)
median2 = median(data2$steps)
sum = sum(data2$steps)-sum(data$steps,na.rm = TRUE)
```

The are `r NAN` values that possess NANs in them. We got around this by approximating the mean value for each interval and replacing each nan with that.

So it looks like the altered mean steps taken per day was `r mean2` and the median was `r median2`. As we can see this had no effect on the mean and median of the data. One thing we do observe is there is a big (`r sum`) increase in the daily estimate of steps.


# Are there differences in activity patterns between weekdays and weekends?

```{r}
library(ggplot2)
Weekdays = weekdays(as.Date(data2$date))
Weekdays  =gsub(c("Saturday|Sunday"),"weekend",Weekdays)
Weekdays  =gsub("Monday|Tuesday|Wednesday|Thursday|Friday","weekday",Weekdays)
data2[,4] = Weekdays
timeseries2 = data2 %>% group_by(interval,V4) %>% summarize(m = mean(steps,na.rm = TRUE))
ggplot(timeseries2,aes(interval,m))+geom_line()+facet_grid(V4~.) + xlab("Interval Duration") +ylab("Steps")
```
